\chapter*{ABSTRACT}
The traditional method of manually screening resumes is a time-consuming and inefficient process that can hinder the recruitment process and lead to missed opportunities to identify top-tier candidates. Additionally, manual screening is susceptible to human biases that can perpetuate inequities in hiring decisions. To address these challenges, this project proposes an innovative approach to resume evaluation and candidate application tracking (ATS) powered by machine learning.

The proposed system utilizes natural language processing (NLP) techniques to extract key information from resumes, including skills, education, and professional experience. This extracted information is then fed into cutting-edge machine learning algorithms that identify and classify skills, rank candidates according to their qualifications and alignment with specific job requirements, and mitigate potential biases in the decision-making process.

This intelligent ATS streamlines the recruitment process by automating the initial screening stage, allowing recruiters to dedicate their time to interacting with the most promising candidates. This leads to enhanced efficiency, reduced time-to-hire, and improved hiring outcomes.

Furthermore, the system promotes fairness and equity in hiring by mitigating potential biases in the decision-making process. This is achieved by employing techniques such as data cleansing, bias-aware algorithms, and diversity metrics. By ensuring that candidates are evaluated objectively and without prejudice, the system fosters a more inclusive and equitable hiring landscape.

The development and implementation of this intelligent ATS can significantly impact organizations by empowering them to identify the most qualified candidates, build high-performing teams, and achieve their strategic goals. By embracing this innovative approach, organizations can transform their recruitment processes, enhance their talent acquisition strategies, and reap the benefits of a more efficient, fair, and inclusive hiring environment.

\textbf{Keywords:} Resume evaluation, Candidate application tracking system (ATS), Machine learning, Natural language processing (NLP), Skill identification, Skill classification, Candidate ranking, Bias mitigation
